%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{preamble.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\hemaClassTitle{}\vspace{-2ex}}

\author[1]{\small Steffen Falgreen\thanks{\texttt{sfl@rn.dk}}}
\author[12]{Anders Ellern Bilgrau\thanks{\texttt{anders.ellern.bilgrau@gmail.com}}}
\author[12]{Lasse Hjort Jakobsen\thanks{\texttt{lasse.j@rn.dk}}}
\author[12]{Jonas Have\thanks{\texttt{jonas.have@rn.dk}}}
\author[12]{Kasper Lindblad Nielsen\thanks{\texttt{k.lindblad@rn.dk}}}
\author[1]{Tarec Christoffer El-Galaly\thanks{\texttt{tarec.galaly@gmail.com}}}
\author[1]{Julie St{\o}ve  B{\o}dker\thanks{\texttt{j.boedker@rn.dk}}}
\author[1]{Alexander Schmitz\thanks{\texttt{alex.schmitz@rn.dk}}}
\author[3]{Ken H. Young\thanks{\texttt{khyoung@mdanderson.org}}}
\author[12]{Hans Erik Johnsen\thanks{\texttt{haej@rn.dk}}}
\author[12]{Karen Dybk{\ae}r\thanks{\texttt{k.dybkaer@rn.dk}}}
\author[12]{Martin B{\o}gsted\thanks{\texttt{martin.boegsted@rn.dk}}}

\affil[1]{Department of Haematology, Aalborg University Hospital}
\affil[2]{Department of Clinical Medicine, Aalborg University}
\affil[3]{Department of Hematopathology, MD Anderson Cancer Center}
\date{\small \today\vspace{-2ex}}


\begin{document}

\maketitle
\textbf{Keywords:~}
Diffuse Large B-Cell Lymphoma; Classification; Chemosensitivity; Gene expression profiling \\

Dozens of genomics based cancer classification systems have been introduced with prognostic, diagnostic, and predictive capabilities.
However, they often employ complex algorithms and are only applicable on whole cohorts of patients, making them difficult to use in a personalized clinical setting.
This prompted us to create \hemaClass{}, available at \url{http://hemaclass.org}, which is an online web application providing an easy interface to one-by-one microarray based risk classifications of diffuse large B-cell lymphoma (DLBCL) into cell-of-origin and chemotherapeutic sensitivity classes.

Users upload patient samples which are RMA normalized and classified depending on settings chosen by the users.
The server accepts samples profiled by the Affymetrix GeneChip HG-U133 Plus 2.0 array.
Three methods of RMA normalization are available: cohort based, reference based, and one-by-one normalization.
The classifications available on \hemaClass{} are:
(1) ABC/GCB which is a classification of DLBCL cases to either activated B-cell phenotype (ABC) or germinal center B-cell phenotype (GCB) \cite{Alizadeh2000},
(2) BAGS which is a refinement of the ABC/GCB classification of DLBCL capable of classifying DLBCL samples into $5$ different B-cell subtypes \cite{DybkaerBoegsted2015},
(3) REGS which is a classification based resistance gene signatures (REGS) for the most prominent drugs used in the treatment of DLBCL patients \cite{Falgreen2015}.
The results are presented on the webserver and are available for download.
An optional text summary of the results and a plot of the predicted progression free survival and overall survival are provided for each patient sample.
The methods are also implemented as an \textsf{R}-package which allows users to run a local instance of the server and thereby avoiding the uploading of large amounts of data.

The server has been running since August 2014 and validated satisfactorily on four different DLBCL cohorts with a total of 882 patients, e.g.\ the ABC/GCB classifications show an approximate $90 \%$ inter-method agreement in all four validation sets.
None of the validation sets have been used in the training of the classifiers.
Since it was initiated, the server has been in routine use in our lab for classification in a number of forthcoming publications, e.g.\ the paper by Marquez et al. \cite{Marquez2015}.

\begingroup
\renewcommand{\section}[2]{}%
\bibliographystyle{plain}
\small
\bibliography{references}
\endgroup
\end{document}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
